---
title: "Data Visualization Group D"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, echo = False, error=FALSE}

#Install packages & Import Libraries

if(!'ggplot2'%in%installed.packages()){
  install.packages('ggplot2')}
if(!'ggthemes'%in%installed.packages()){
  install.packages('ggthemes')}
if(!'nlme'%in%installed.packages()){
  install.packages('nlme')}
if(!'gapminder'%in%installed.packages()){
  install.packages('gapminder')}
if(!'gganimate'%in%installed.packages()){
  install.packages('gganimate')}
if(!'ggExtra'%in%installed.packages()){
  install.packages('ggExtra')}
if(!'psych'%in%installed.packages()){
  install.packages('psych')}
if(!'reshape2'%in%installed.packages()){
  install.packages('reshape2')}
if(!'dplyr'%in%installed.packages()){
  install.packages('dplyr')}
if(!'nycflights13'%in%installed.packages()){
  install.packages('nycflights13')}
if(!'ggcorrplot'%in%installed.packages()){
  install.packages('ggcorrplot')}
if(!'waffle'%in%installed.packages()){
  install.packages('waffle')}
if(!'tidyr'%in%installed.packages()){
  install.packages('tidyr')}
if(!'scales'%in%installed.packages()){
  install.packages('scales')}
if(!'ggalt'%in%installed.packages()){
  install.packages('ggalt')}
if(!'data.table'%in%installed.packages()){
  install.packages('data.table')}
if(!'data.table'%in%installed.packages()){
  install.packages('data.table')}
if(!'data.table'%in%installed.packages()){
  install.packages('data.table')}
if(!'extrafont'%in%installed.packages()){
  install.packages('extrafont')}
if(!'lubridate'%in%installed.packages()){
  install.packages('lubridate')}
if(!'DT'%in%installed.packages()){
  install.packages('DT')}
if(!'grid'%in%installed.packages()){
  install.packages('grid')}
if(!'gridExtra'%in%installed.packages()){
  install.packages('gridExtra')}
if(!"prettydoc" %in% installed.packages()) {
  install.packages("prettydoc")}
if(!"devtools" %in% installed.packages()) {
  install.packages("devtools")}
if(!"tidyverse" %in% installed.packages()) {
  install.packages("tidyverse")}
if(!"ggdark" %in% installed.packages()) {
  install.packages("ggdark")}
if(!"here" %in% installed.packages()) {
  install.packages("here")}
if(!"png" %in% installed.packages()) {
  install.packages("png")}
if(!"gifski" %in% installed.packages()) {
  install.packages("gifski")}
if(!"data.table" %in% installed.packages()) {
  install.packages("data.table")}
if(!"forcats" %in% installed.packages()) {
  install.packages("forcats")}
if(!"tufte" %in% installed.packages()) {
  install.packages("tufte")}
if(!"colorspace" %in% installed.packages()) {
  install.packages("colorspace")}
if(!"viridisLite" %in% installed.packages()) {
  install.packages("viridisLite")}
if(!"formatR" %in% installed.packages()) {
  install.packages("formatR")}
if(!"DiagrammeR" %in% installed.packages()) {
  install.packages("DiagrammeR")}
if(!"xaringan" %in% installed.packages()) {
  install.packages("xaringan")}
if(!"ggridges" %in% installed.packages()) {
  install.packages("ggridges")}
if(!"GGally" %in% installed.packages()) {
  install.packages("GGally")}
if(!"kableExtra" %in% installed.packages()) {
  install.packages("kableExtra")}

# Connect with the libaries
library(ggplot2)
library(ggthemes)
library(nlme)
library(gganimate)
library(gapminder)
library(ggExtra)
library(psych)
library(reshape2)
library(dplyr)
library(nycflights13)
library(ggcorrplot)
library(waffle)
library(tidyr)
library(scales)
library(ggalt)
library(data.table)
library(extrafont)
library(lubridate)
library(DT)
library(grid)
library(gridExtra)
library(prettydoc)
library(devtools)
library(tidyverse)
library(ggdark)
library(here)
library(png)
library(gifski)
library(forcats)
library(tufte)
library(colorspace)
library(viridisLite)
# library(Zelig)
library(formatR)
library(DiagrammeR)
library(xaringan)
library(ggridges)
library(GGally)
library(xlsx)
library(scales)
library(kableExtra)

```



```{r, echo = FALSE, error = FALSE}

#Defining the general colors to avoid hard coding  
fill_color = 'white'
decoration_color = '#778287'
main1_color = '#3366ff'
main2_color = '#FFCC33'

#Creare a theme we can apply everywhere 

groupD_theme<-theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  plot.title = element_text(size = 14, hjust = 0.2, color = decoration_color),
  axis.title = element_text(size = 10, hjust = 0.5, color = decoration_color),
  axis.text = element_text(colour = decoration_color, size = 8),
  axis.ticks = element_blank(),
  axis.line = element_line(colour = decoration_color, size=0.3, linetype = "dashed"), 
  panel.border = element_blank(),
  panel.grid = element_blank(),
  strip.text = element_text(size = 12, color = decoration_color),
  panel.background = element_blank(),
  strip.background =element_blank(),
  plot.background = element_blank(),
  legend.position="None"
)

#Now we set the new defined theme to the default option 
theme_set(groupD_theme)

#Creare a theme that includes a legend 
groupD_theme2<-theme_bw() + theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  plot.title = element_text(size = 14, hjust = 0.2, color = decoration_color),
  axis.title = element_text(size = 10, hjust = 0.5, color = decoration_color),
  axis.text = element_text(colour = decoration_color, size = 8),
  axis.ticks = element_blank(),
  axis.line = element_line(colour = decoration_color, size=0.3, linetype = "dashed"), 
  panel.border = element_blank(),
  panel.grid = element_blank(),
  strip.text = element_text(size = 12, color = decoration_color),
  panel.background = element_blank(),
  strip.background =element_blank(),
  plot.background = element_blank(),
  legend.text	= element_text(size = 10, hjust = 0.5, color = decoration_color), 
  legend.position = c(0.815, 0.27),
  legend.key = element_blank(),
  legend.title = element_blank()
)

theme_set(groupD_theme2)

```


##### Our group chose to visualise data on Youth Unemployment in Latin America & The Caribbean. This dataset contains statistics on employment by industry, country, gender and service type. Our visual encoding is centered around the question: What are the key characteristics of Youth Unemployment in Latin America? To understand the problem, let us carry out some exploratory analysis. 


```{r dataset_view, echo = FALSE, include = FALSE}

# JD Path: path_data <- "/Users/juandavidbotero/Desktop/Data Visualization/Youth_Latam.csv"
# AS Path: 

path_data <- "C:/Users/anups/Google Drive/Group D/Data Visualization/Youth_Latam.csv"

youth_latam <- data.table(read.csv(path_data, header = T))
colnames(youth_latam) <- c("Country", colnames(youth_latam[,2:length(youth_latam)]))

#We remove all rows with NA's

youth_latam <- na.omit(youth_latam)


```


### We first calculated the Unemployment Rates in the countries for which we had enough data. The rates were then plotted as follows:


```{r unemployment_rates, echo = FALSE}


###################################################### UNEMPLOYMENT RATES ####################################################

################# A PLOT OF THE UNEMPLOYMENT RATES BY COUNTRY ###########################

############################# Unemployment_Rate_Calculations


### Steps to create the Unemployed_Youth dataset:
    ### Sum the Economically Active Population by country
    ### Sum the Unemployed population country
    ### Division of the two gives us the Unemployment Rate

Unemployed_Youth <- youth_latam[,list(Economically_Active = sum(Economically.Active.Population), 
                                      Unemployed_Population = sum(Unemployed)), by = "Country"][
                                        , list(Unemployment_Rate = (Unemployed_Population / Economically_Active)*100),
                                        by = "Country"]

Unemployed_Youth <-Unemployed_Youth[order(Unemployed_Youth$Unemployment_Rate), ]  # sort the dataset by the employment rate


Unemployed_Youth$Country <- factor(Unemployed_Youth$Country, levels = Unemployed_Youth$Country) #to keep the factor order as the unemployment



############################# Graphing Unemployment_Rates

theme_set(groupD_theme) # set theme


Unemployment_Lollipop <- ggplot(Unemployed_Youth, aes(x=Country, y=Unemployment_Rate)) + 
  geom_point(col= main1_color, 
             size= 3, 
             shape = 19) +   # Draw points
  geom_segment(aes(x=Country, 
                   xend=Country, 
                   y=0, 
                   yend=Unemployment_Rate), 
               color = main2_color,
               linetype = 'dotted') + # Draw dashed line
  labs(x = NULL, y= NULL, caption = "Unemployment Rate (%)") +
    theme(axis.text.x = element_text(angle=65, vjust= 0.4))


print(Unemployment_Lollipop, position = c(0.5, 0, 1, 1))
```
# This plot shows the unemployement rate per country in a lollipop stile. This way the focus is more in the rate, rather than the coutries. It's easier to show how the unemployement increases.


#####TUFTE PLOT 
# After the exploratory analysis we continue with the tufte chart, we have decided to use a boxplot to show the distribution of the employement per industries in LATAM. First we separate into two sections, top umemployed countries and bottom umeployed countries. Then we do a box plot of the top and bottom countries. 

```{r unemployment_rates, echo = FALSE}

############################################## TUFTE PLOT ######################################

################# A PLOT OF THE NUMBER OF PEOPLE IN EACH POSITION IN THE TWO HALFS (MEDIAN AND SPREAD) ###########################

############################# Subsetting the main dataset for those countries that are above the average Unemployment Rate and those
############################# that are below

top_unemployed <- youth_latam[Country %in% (Unemployed_Youth[Unemployed_Youth$Unemployment_Rate >= mean(Unemployed_Youth$Unemployment_Rate)]$Country)]


bottom_unemployed <- youth_latam[Country %in% (Unemployed_Youth[Unemployed_Youth$Unemployment_Rate < mean(Unemployed_Youth$Unemployment_Rate)]$Country)]


############################# Tufte Plot for the bottom half


### Steps to create the Bottom_Type dataset:
    ###  Subset the main bottom_unemployed dataset for only the required columns (position)
    ###  Sum the positions, also putting them as ccolumns and group by coutntry

Bottom_Type <- bottom_unemployed[,bottom_unemployed[,c(1,22:28)]][,list(Type = colnames(.SD), Employed = sapply(.SD, sum)),
                                                         by = "Country"]


### Graph representing the number of people employed by Position

theme_set(groupD_theme2)

Bottom_Positions_Graph <- ggplot(Bottom_Type, aes(Type, Employed / 1000)) +
geom_tufteboxplot(size = 1, whisker.type = 'line', color = main1_color) + 
      theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
      labs(x = NULL, y  = NULL)



############################# Tufte Plot for the Top half


### Steps to create the Top_Type dataset same as for the bottom type

Top_Type <- top_unemployed[,top_unemployed[,c(1,22:28)]][,list(Type = colnames(.SD), Employed = sapply(.SD, sum)), by = "Country"]

                                                         
### Graph representing the number of people employed by Position                                                      
                                                                                                            
theme_set(groupD_theme2)

Top_Positions_Graph <- ggplot(Top_Type, aes(Type, Employed / 1000)) +
geom_tufteboxplot(size = 1, whisker.type = 'line', color = main1_color) + 
      theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
      labs(caption = "Employment Industry", x = NULL, y  = NULL)


grid.arrange(Bottom_Positions_Graph, Top_Positions_Graph, ncol = 2)


```

# Seen both plots we can tell there are some differences, in the countries where the unemployement is higher the employement in the public sector is higher (Services, Armed Forces, Administrative),on the other hand they are lower in countries with lower unemployement and the mean of the sectors stays the same but the outliers are lower.


```{r}

############################################################ POPULATION PLOT ##############################

################# A PLOT OF THE PROPORTION OF MEN AND WOMEN IN EACH POSITON OF THE BOTTOM AND TOP HALFS ###########################


############################# Reinitialise the top and bottom halfs of the main dataset

top_unemployed <- youth_latam[Country %in% (Unemployed_Youth[Unemployed_Youth$Unemployment_Rate >= mean(Unemployed_Youth$Unemployment_Rate)]$Country)]


bottom_unemployed <- youth_latam[Country %in% (Unemployed_Youth[Unemployed_Youth$Unemployment_Rate < mean(Unemployed_Youth$Unemployment_Rate)]$Country)]


############################# Bottom_Type_Pr dataset

### Steps to create the Bottom_Type_Pr dataset:
    ###  Subset the main bottom_unemployed dataset for only the required columns (position)
    ###  Sum the positions, also putting them as ccolumns, divide by the sum of the total positions and group by gender 


Bottom_Type_Pr <- bottom_unemployed[,bottom_unemployed[,c(2,30,22:28)]][,list(Type = colnames(.SD), 
                                                                          Employed_Proportion = (sapply(.SD, sum)/ sum(Total.Employment.Type)
                                                                                                 *100)),by = c("Demographic")]

Bottom_Type_Pr <- Bottom_Type_Pr[Bottom_Type_Pr[,Bottom_Type_Pr$Type != "Total.Employment.Type"],]  #remove the total (100%) 


############################# Bottom_Type_Pr Plot


### Steps to create the Bottom_Type_Pr plot:
    ###  Prepare the Men column as -ve to be able to do a population plot 
    ### Set breaks and labels
    ### Plot


theme_set(groupD_theme2)

Bottom_Type_Pr[Bottom_Type_Pr$Demographic == "Men",][,"Employed_Proportion"] <- Bottom_Type_Pr[Bottom_Type_Pr$Demographic ==
                                                                                                     "Men",][,"Employed_Proportion"]*-1

brks <- seq(-50, 50, 5)
lbls <- paste0(as.character(c(seq(50, 0, -5), seq(5, 50, 5))))

Bottom_Unemployed_Youth_Graph <- ggplot(Bottom_Type_Pr, aes(x = Type, y = Employed_Proportion, fill = Demographic)) +   # Fill column
 geom_bar(stat = "identity", width = .6) +   # draw the bars
 scale_y_continuous(breaks = brks,   # Breaks
                    labels = lbls) + # Labels
 coord_flip() +  # Flip axes  
 scale_fill_brewer(palette = "Paired") +
  labs (x =  NULL, y = NULL)



############################# Top_Type_Pr plot


### Steps to create the Top_Type_Pr plot same as bottom_type_pr


Top_Type_Pr <- top_unemployed[,top_unemployed[,c(2,30,22:28)]][,list(Type = colnames(.SD), 
                                                                          Employed_Proportion = (sapply(.SD, sum)/ sum(Total.Employment.Type)
                                                                                                 *100)), by = c("Demographic")]


Top_Type_Pr <- Top_Type_Pr[Top_Type_Pr[,Top_Type_Pr$Type != "Total.Employment.Type"],] # remove 100% row


############################# Top_Type_Pr Plot


### Steps to create the Top_Type_Pr plot same as for the Bottom_Type_Pr plot


Top_Type_Pr[Top_Type_Pr$Demographic == "Men",][,"Employed_Proportion"] <- Top_Type_Pr[Top_Type_Pr$Demographic ==
                                                                                                     "Men",][,"Employed_Proportion"]*-1

brks <- seq(-50, 50, 5)
lbls <- paste0(as.character(c(seq(50, 0, -5), seq(5, 50, 5))))


Top_Unemployed_Youth_Graph <- ggplot(Top_Type_Pr, aes(x = Type, y = Employed_Proportion, fill = Demographic)) +   # Fill column
 geom_bar(stat = "identity", width = .6) +   # draw the bars
 scale_y_continuous(breaks = brks,   # Breaks
                    labels = lbls) + # Labels
 coord_flip() +  # Flip axes
 scale_fill_brewer(palette = "Paired")+
  xlab("") +
  ylab ("")+
  labs (caption = "% Employed", title = "Above Average")


############################## Print 

grid.arrange(Bottom_Unemployed_Youth_Graph, Top_Unemployed_Youth_Graph, ncol = 2)


```

# For advance DataViz we have choosen to do first a Dumbbel Plot. This plot is great for displaying differences between two groups. 

```{r}

####################################################### DUMBELL PLOT ###########################################


################# DISPERTION OF EMPLOYMENT RATE AMONGST MEN AND WOMEN ###########################


############################# Reinitialize the top and bottom half

top_unemployed <- youth_latam[Country %in% (Unemployed_Youth[Unemployed_Youth$Unemployment_Rate >= mean(Unemployed_Youth$Unemployment_Rate)]$Country)]


bottom_unemployed <- youth_latam[Country %in% (Unemployed_Youth[Unemployed_Youth$Unemployment_Rate < mean(Unemployed_Youth$Unemployment_Rate)]$Country)]


############################# Bottom_Unemployed_Youth 

### Steps to create the Bottom_Unemployed_Youth dataset are the same as that for Unemployment_Rates

Bottom_Unemployed_Youth <- bottom_unemployed[,list(Demographic, Economically_Active = sum(Economically.Active.Population),
                                      Unemployed_Population = sum(Unemployed)), by = c("Country","Demographic")][
                                        ,list(Demographic, Unemployment_Rate = Unemployed_Population / Economically_Active),
                                        by = c("Country","Demographic")][,list(Country, Demographic , Unemployment_Rate = (Unemployment_Rate *
                                                                                                                             100))]

### Additional Steps:
    ### Filter the Bottom_Unemployed_Youth Dataset for the men and women, separate them into datasets
    ### Merge the newly created dataset to have a single one that can be used for plotting  
    ### Change the Column Names for easy subsetting
    ### Sort the values in descending order and factor the categorical variables to keep the order

Bottom_Unemployed_Youth_Men <- Bottom_Unemployed_Youth[Bottom_Unemployed_Youth$Demographic == "Men", c("Country","Unemployment_Rate")]
Bottom_Unemployed_Youth_Women <- Bottom_Unemployed_Youth[Bottom_Unemployed_Youth$Demographic == "Women", c("Country","Unemployment_Rate")]


Bottom_Unemployed_Youth_Gender <- merge(Bottom_Unemployed_Youth_Men, Bottom_Unemployed_Youth_Women, by ="Country")
colnames(Bottom_Unemployed_Youth_Gender)  <- c("Country","Unemployment_Rate_Men", "Unemployment_Rate_Women")


Bottom_Unemployed_Youth_Gender <- Bottom_Unemployed_Youth_Gender[order(Bottom_Unemployed_Youth_Gender$Unemployment_Rate_Women), ]  # sort
Bottom_Unemployed_Youth_Gender$Country <- factor(Bottom_Unemployed_Youth_Gender$Country, levels = Bottom_Unemployed_Youth_Gender$Country)


### Steps to create the Bottom_Gender_Plot plot:
    ###  Create Dumbell plot that highlights the gender difference

Bottom_Gender_Plot <- ggplot(Bottom_Unemployed_Youth_Gender, aes(x=Unemployment_Rate_Men, xend=Unemployment_Rate_Women, y=Country)) + 
        geom_dumbbell(colour_xend = main1_color,
                      colour_x = main2_color,
                      size= 1,
                      color = main2_color) + 
        scale_x_continuous() +
  labs(x=NULL, 
     y=NULL, 
     caption="Unemployment Rates Bottom Half (%)")


############################# Top_Unemployed_Youth 


### Steps to create the Top_Gender_Plot are the same as the one for the Bottom Gender Plot


Top_Unemployed_Youth <- top_unemployed[,list(Demographic, Economically_Active = sum(Economically.Active.Population),
                                      Unemployed_Population = sum(Unemployed)), by = c("Country","Demographic")][
                                        ,list(Demographic, Unemployment_Rate = Unemployed_Population / Economically_Active),
                                        by = c("Country","Demographic")][,list(Country, Demographic , Unemployment_Rate = (Unemployment_Rate * 100))]



### Additional Steps


Top_Unemployed_Youth_Men <- Top_Unemployed_Youth[Top_Unemployed_Youth$Demographic == "Men", c("Country","Unemployment_Rate")]
Top_Unemployed_Youth_Women <- Top_Unemployed_Youth[Top_Unemployed_Youth$Demographic == "Women", c("Country","Unemployment_Rate")]


Top_Unemployed_Youth_Gender <- merge(Top_Unemployed_Youth_Men, Top_Unemployed_Youth_Women, by ="Country")
colnames(Top_Unemployed_Youth_Gender)  <- c("Country","Unemployment_Rate_Men", "Unemployment_Rate_Women")

Top_Unemployed_Youth_Gender <- Top_Unemployed_Youth_Gender[order(Top_Unemployed_Youth_Gender$Unemployment_Rate_Women), ]  # sort
Top_Unemployed_Youth_Gender$Country <- factor(Top_Unemployed_Youth_Gender$Country, levels = Top_Unemployed_Youth_Gender$Country)



### Plotting



Top_Gender_Plot <- ggplot(Top_Unemployed_Youth_Gender, aes(x=Unemployment_Rate_Men, xend=Unemployment_Rate_Women, y=Country)) + 
        geom_dumbbell(colour_xend = main1_color,
                      colour_x = main2_color,
                      size = 1,
                      colour = main2_color) + 
        scale_x_continuous() +
  labs(x=NULL, 
     y=NULL, 
     caption="Unemployment Rates Top Half (%)")
    

####

grid.arrange(Bottom_Gender_Plot,Top_Gender_Plot, ncol = 2)


```


# As we can see, the plot shows the unemployement rates in all the countries and shows the difference between male and female. As we can see, in the Dominican Republic is where the biggest difference between men and women is. The female unemployement is near 25% whereas the male unemployement is close to 5%. On the other side, Peru has the smallest difference of unemployement between genders. Also Mexico and Peru show the oposite tendency from the other countries, in these both countries the femele umeployement is lower than the male unemployement 


```{r}


############################### WAFFLE CHART  ################################# 




```


